---
title: ""
author: "Vandrade"
date: "2024/13/06"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Importing librarys

```{r importing librarys, warning=FALSE, message=FALSE}

library(tidyverse)
library(electionsBR)
library(sf)
library(geobr)
library(extrafont)
library(foreign)
library(rdrobust)
library(readxl)
library(microdatasus)
library(gt)


```

# electionsBR

```{r}
year = c(2000, 2004, 2008, 2012, 2016, 2020)

```

## VOTE_MUN_ZONE

```{r votes per municipality's zones, cache = TRUE}


fun_vote_mun_zone <- function(year) {
  elections_tse(year, type = 'vote_mun_zone', uf = 'all')}

df_vote_mun_zone <- map_df(year, fun_vote_mun_zone)

df_candidates_votes <- df_vote_mun_zone |> 
  mutate(votos_zona = coalesce(QT_VOTOS_NOMINAIS_VALIDOS, QT_VOTOS_NOMINAIS)) |> 
  select(ano = ANO_ELEICAO,
         estado = SG_UF,
         cd_municipio_tse = CD_MUNICIPIO,
         nm_municipio = NM_MUNICIPIO,
         zona = NR_ZONA,
         cargo = DS_CARGO,
         nr_candidato = NR_CANDIDATO,
         nm_candidato = NM_CANDIDATO,
         votos_zona,
         partido = SG_PARTIDO) |> 
  filter(cargo == 'Prefeito') |> 
  group_by(ano, estado, cd_municipio_tse, nr_candidato) |> 
  summarize(nm_municipio = first(nm_municipio),
            nm_candidato = first(nm_candidato),
            total_votos = sum(votos_zona),
            partido = first(partido)) |> 
  group_by(ano, estado, cd_municipio_tse) |> # adding a column "result". 
  mutate(resultado = ifelse(total_votos == max(total_votos), 'Eleito', 'Não Eleito')) |> 
  mutate(cd_municipio_tse = as.numeric(cd_municipio_tse))

df_candidates_votes |> head(3)


```

## DETAILS_MUN_ZONE

this dataframe is at zone level, not candidate. Its very important to
have it because it will be used to filter the municipalities with up to
200000 electoral population (municipalities without 2nd turn)

```{r, cache = TRUE}

fun_details_mun_zone <- function(year) {
  elections_tse(year, type = 'details_mun_zone', uf = 'all')}

df_details_mun_zone <- map_df(year, fun_details_mun_zone)

df_pop_votante <- df_details_mun_zone |> 
  select(ano = ANO_ELEICAO,
         estado = SG_UF,
         cd_municipio_tse = CD_MUNICIPIO,
         nm_municipio = NM_MUNICIPIO,
         zona = NR_ZONA,
         cargo = DS_CARGO,
         pop_votante = QT_APTOS,
         nr_abstencoes = QT_ABSTENCOES) |> 
  filter(cargo == 'Prefeito') |> 
  group_by(ano, estado, cd_municipio_tse) |> 
  summarize(nm_municipio = first(nm_municipio),
            pop_votante = sum(pop_votante)) |> 
  mutate(cd_municipio_tse = as.numeric(cd_municipio_tse))

df_pop_votante |> head(3)



```



## CANDIDATE_LOCAL

```{r, cache=TRUE}

fun_candidates <- function(year) {
  candidate_local(year, uf = 'all') |> 
    mutate(NR_PROTOCOLO_CANDIDATURA = as.character(NR_PROTOCOLO_CANDIDATURA),
           NR_CPF_CANDIDATO = as.character(NR_CPF_CANDIDATO),
           CD_COR_RACA = as.character(CD_COR_RACA))}

df_fun_candidates <- map_df(year, fun_candidates)

df_candidates_details <- df_fun_candidates |> 
  select(ano = ANO_ELEICAO, #selecionando e filtrando as variaveis
         estado = SG_UF,
         cargo = DS_CARGO,
         nr_candidato = NR_CANDIDATO,
         nm_candidato = NM_CANDIDATO,
         partido = SG_PARTIDO,
         idade = NR_IDADE_DATA_POSSE,
         genero = DS_GENERO,
         educacao = DS_GRAU_INSTRUCAO,
         cor_raca = DS_COR_RACA,
         ocupacao = DS_OCUPACAO) |> 
  filter(cargo == 'PREFEITO') |> 
  group_by(ano, estado, partido, nr_candidato, nm_candidato) |> # some of the candidates are repeated. 
  summarise(idade = max(idade),
            genero = first(genero),
            educacao = first(educacao),
            cor_raca = first(cor_raca),
            ocupacao = first(ocupacao))
# this dataframe had many errors.

df_candidates_details |> head(3)

```

## creating a complete candidate dataframe.

With personal characteristics, votes recieved, IBGE's municipality code
and electoral population

what lacks here is municipality information. the most important
information we should have here is the IBGE municipality code. With this
code we`ll be able to collect and implement complementary information
from most sources.

### IBGE CODE

```{r creating complete candidates dataframe, cache=TRUE}

cd_ibge <- read_csv("Dados/municipios_brasileiros_tse.csv") |> 
  select(codigo_tse, codigo_ibge) |>           
  mutate(codigo_ibge_sus = as.character(codigo_ibge),
         codigo_ibge_sus = substring(codigo_ibge_sus, 1, 6),
         codigo_ibge_sus = as.numeric(codigo_ibge_sus))

```

### GEOBR

A first example of data that can be implemented throug IBGE's
municipality code is the geometry of each municipality

```{r GEOBR adicionando geometrias, cache = TRUE}

muni_geomentry <- read_municipality(code_muni = 'all') |> 
  select(code_muni, geom)

```

### POPULATION

```{r}

pop_censo <- read_xlsx("Dados/pop-sexo-censo-2010.xlsx") |> 
  rename(cd_muni = Município,
         pop_mas = Masculino, 
         pop_fem = Feminino) |> 
  mutate(cd_muni = substr(cd_muni,1, 6),
         cd_muni = as.numeric(cd_muni),
         pop_total = pop_mas + pop_fem,
         prop_fem_pop = pop_fem/pop_total)

```

## CANDIDATES - DataFrame

```{r creating complete candidates dataframe}

candidates <- df_candidates_votes |> 
  left_join(df_candidates_details, by = c('ano', 'estado', 'nr_candidato', 'nm_candidato', 'partido')) |> 
  left_join(cd_ibge, by= c('cd_municipio_tse' = 'codigo_tse')) |> 
  left_join(muni_geomentry, by=c('codigo_ibge' = 'code_muni')) |> 
  left_join(df_pop_votante, by= c('ano', 'estado', 'cd_municipio_tse', 'nm_municipio')) |>
  left_join(pop_censo, by = c('codigo_ibge_sus' = 'cd_muni')) |> 
  mutate(genero = ifelse(genero %in% c("NÃO INFORMADO", "NÃO DIVULGÁVEL"), NA, genero))


```



```{r}
# saveRDS(candidates, "candidates.rds")
# 
candidates <- readRDS('candidates.rds')

```


```{r visual - gender of all mayors elected}

candidates |>
  filter(resultado == 'Eleito') |> 
  ggplot(aes(fill = genero)) + 
  geom_sf(aes(geometry = geom), color = NA) +
  facet_wrap(~ ano, ncol = 3) + 
  theme_void() +
  labs(title = "Distribuição de Candidatos Eleitos por Gênero",
       subtitle = "Eleições de 2000 a 2020",
       fill = "") + 
  scale_fill_manual(values = c("#ff7373", "#7373ff", 'white'),
                    labels = c("Mulher Eleita", "Homem Eleito", 'Não Informado')) +
  labs(caption = "Por InsperData") +
  theme(
    legend.position = "bottom",
    text=element_text(family="Segoe UI"),
    plot.subtitle = element_text(margin = margin(b = 6)))



```


## Creating Male x Female candidates dataframe

```{r}
mayors <- candidates |> 
  group_by(ano, estado, cd_municipio_tse) |> 
  arrange(desc(total_votos)) |> 
  slice_head(n = 2) |> 
  filter(!is.na(genero)) |> 
  group_by(ano, estado, cd_municipio_tse) |> 
  filter(n_distinct(genero) == 2) |> 
  mutate(mg_fem = (total_votos[genero == 'FEMININO'] - total_votos[genero == 'MASCULINO'])/sum(total_votos),
         pop_total = sum(pop_mas + pop_fem)) |>
  filter(resultado == 'Eleito', pop_votante <= 200000)


```


# SINAN 2008 - 2022

path_dbf = 'dados/sinan_dbf'

list <- list.files(path = path_dbf, pattern = "\\.dbf$", full.names = TRUE) 

list_of_dfs <- lapply(list, read.dbf)

sinan_raw <- do.call(rbind, list_of_dfs)

saveRDS(sinan_raw, "sinan_raw.rds")

```{r}

sinan_raw <- readRDS('sinan_raw.rds')

```



```{r, cache=TRUE}

sinan_df <- sinan_raw |> 
  select(
    ano_nasc = ANO_NASC,
    tp_lesao = CIRC_LESAO, #(x85 - y09)
    ano = NU_ANO,
    cd_uf_not = SG_UF_NOT,
    cd_muni_not = ID_MUNICIP,
    idade = NU_IDADE_N,
    sexo = CS_SEXO,
    gestante = CS_GESTANT,
    raca = CS_RACA,
    escolaridade = CS_ESCOL_N,
    cd_uf_res = SG_UF,
    cd_muni_res = ID_MN_RESI,
    conjugal = SIT_CONJUG,
    cd_uf_ocor = SG_UF_OCOR,
    cd_muni_ocor = ID_MN_OCOR,
    hora_ocor = HORA_OCOR,
    local_ocor = LOCAL_OCOR,
    outras_vezes = OUT_VEZES,
    v_autop = LES_AUTOP,
    motivo_viol = VIOL_MOTIV, # (1 = sexismo, 2 = homofobia, 3 = racismo)
    v_fisica = VIOL_FISIC, # (1 = sim, 2 = nao) 
    v_psico = VIOL_PSICO,
    v_infantil = VIOL_INFAN,
    v_tort = VIOL_TORT,
    v_sexu = VIOL_SEXU,
    v_finan = VIOL_FINAN,
    v_negli = VIOL_NEGLI,
    agressor_sexo = AUTOR_SEXO,
    evolucao = EVOLUCAO) |> 
  mutate(tp_lesao = as.character(tp_lesao),
         idade_ofc = as.numeric(as.character(ano))  - as.numeric(as.character(ano_nasc))) |> 
  select(idade_ofc, ano_nasc, ano, everything())|> 
  filter(idade_ofc >= 15 & idade_ofc <= 49)


```



## Agrupando dados para MANDATO



```{r}


notificacoes_df <- sinan_df |>
  mutate(tp_lesao = as.factor(tp_lesao),
         cd_muni_res = as.numeric(as.character(cd_muni_res)),
         ano = as.numeric(as.character(ano))) |> 
  group_by(ano, cd_muni_res) |> 
  summarise(noti_total = n(),
            noti_autop = sum(v_autop == 1),
            noti_fem_autop = sum(v_autop == 1 & sexo == 'F'),
            noti_fem = sum(sexo == 'F'),
            noti_fisica = sum(v_fisica == 1),
            noti_fem_fisica = sum(v_fisica == 1 & sexo == 'F'),            
            noti_psico = sum(v_psico == 1),
            noti_fem_psico = sum(v_psico == 1 & sexo == 'F'),
            noti_infantil = sum(v_infantil == 1),
            noti_fem_infantil = sum(v_infantil == 1 & sexo == 'F'),
            noti_tort = sum(v_tort == 1),
            noti_fem_tort = sum(v_tort == 1 & sexo == 'F'),
            noti_sexu = sum(v_sexu == 1),
            noti_fem_sexu = sum(v_sexu == 1 & sexo == 'F'),
            noti_finan = sum(v_finan == 1),
            noti_fem_finan = sum(v_finan == 1 & sexo == 'F'),
            noti_negli = sum(v_negli == 1),
            noti_fem_negli = sum(v_negli == 1 & sexo == 'F')) |>  
  mutate(mandato = case_when( 
    ano %% 4 == 0 ~ ano - 4,
    ano %% 4 == 1 ~ ano - 1,
    ano %% 4 == 2 ~ ano - 2,
    ano %% 4 == 3 ~ ano - 3)) |>
  group_by(mandato, cd_muni_res) |> 
  summarise(noti_total_mandato = sum(noti_total),
            noti_fem_mandato = sum(noti_fem),
            noti_autop_mandato = sum(noti_autop),
            noti_fem_autop_mandato = sum(noti_fem_autop),
            noti_fisica_mandato = sum(noti_fisica),
            noti_fem_fisica_mandato = sum(noti_fem_fisica),
            noti_psico_mandato = sum(noti_psico),
            noti_fem_psico_mandato = sum(noti_fem_psico),
            noti_infantil_mandato = sum(noti_infantil),
            noti_fem_infantil_mandato = sum(noti_fem_infantil),
            noti_tort_mandato = sum(noti_tort),
            noti_fem_tort_mandato = sum(noti_fem_tort),
            noti_sexu_mandato = sum(noti_sexu),
            noti_fem_sexu_mandato = sum(noti_fem_sexu),
            noti_finan_mandato = sum(noti_finan),
            noti_fem_finan_mandato = sum(noti_fem_finan),
            noti_negli_mandato = sum(noti_negli),
            noti_fem_negli_mandato = sum(noti_fem_negli)) |> 
  mutate(mutate(across(starts_with("noti_"), ~ replace_na(., 0))))

```


# Juntando dados do tse com dados do sinan


```{r}

df_terms <- mayors |> 
  left_join(notificacoes_df, by = c('ano' = 'mandato', 'codigo_ibge_sus' = 'cd_muni_res')) |> 
  filter(ano >= 2008 & ano <= 2020,
         nm_candidato != 'JAILSON DA COSTA ALVES') |> 
  mutate(mulher_eleita = ifelse(mg_fem > 0, TRUE, FALSE),
       across(starts_with('noti_') & ends_with('_mandato'), 
              ~ (./pop_total)*100000),
       across(starts_with('noti_') & ends_with('_mandato'), 
              ~ replace_na(., 0)),
       cd_municipio_tse = as.factor(cd_municipio_tse),
       ano = as.factor(ano))
  
```


### HISTOGRAMA - distribuição da margem das eleicoes
```{r}

df_terms |> 
  ggplot(aes(x = mg_fem, fill = mulher_eleita)) + 
  geom_histogram(boundary = 0, binwidth = 0.05, color = 'white') + 
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = c("#7373ff", "#ff7373"),
                    labels = c("Homem Eleito", "Mulher Eleita")) +
  labs(title = "Distribuição de Margens de Vitória em Municípios Brasileiros",
       subtitle = "Eleições entre Homens e Mulheres (2008 - 2016)",
       x = "Margem de Vitória de Mulheres",
       y = NULL,
       color = 'Woman Elected',
       fill = NULL) + 
  theme_minimal() + 
  theme(legend.position = "top",
        plot.caption = element_text(color = "#3494ba", size = 11),
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 11, color = 'black'),
        axis.title.y = element_text(margin = margin(r = 10, unit = "pt"), size = 11), 
        axis.title.x = element_text(margin = margin(t = 10, unit = "pt"), size = 11),
        text=element_text(family="Segoe UI", size=12)) +
  scale_x_continuous(labels = scales::percent)

```


# Identification Hypothesis: discontinuity

```{r}

rdplot(df_terms$noti_fem_autop_mandato, df_terms$mg_fem, c = 0, nbins = c(100, 100), kernel = 'triangular', x.lim = c(-.50,.5), y.lim = c(0, 150), p=1)
  
rdplot(df_terms$tx_noti_fem_mandato, df_terms$mg_fem, c = 0, nbins = c(100, 100), kernel = 'triangular', x.lim = c(-.20,.20), y.lim = c(0, 150))

rdplot(df_terms$tx_noti_fisica_mandato, df_terms$mg_fem, c = 0, nbins = c(100, 100), kernel = 'triangular', x.lim = c(-.20,.20), y.lim = c(0,100))
  
rdplot(df_terms$tx_noti_fem_fisica_mandato, df_terms$mg_fem, c = 0, nbins = c(100, 100), kernel = 'triangular', x.lim = c(-.20,.20), y.lim = c(0, 100))

rdplot(df_terms$noti_psico_mandato, df_terms$mg_fem, c = 0, nbins = c(50, 50), kernel = 'uniform', y.lim = c(0, 50), x.lim = c(-.20,.20), p=1)

rdplot(df_terms$noti_fem_psico_mandato, df_terms$mg_fem, c = 0, nbins = c(100, 100), kernel = 'triangular', x.lim = c(-.20,.20), y.lim = c(0, 50), p=1)


```


```{r}

controls <- model.matrix(~ pop_total + prop_fem_pop + ano, data = df_terms)

rd_robust_result <- rdrobust(df_terms$noti_total_mandato, df_terms$mg_fem,
                               covs = controls,
                               c = 0,
                               kernel = 'uniform',
                               cluster = df_terms$cd_municipio_tse)

summary(rd_robust_result)


```

```{r}

controls <- model.matrix(~ pop_total + prop_fem_pop + ano + partido + educacao, data = df_terms)

rdrobust_result <- rdrobust(df_terms$noti_autop_mandato, df_terms$mg_fem,
                              covs = controls,
                              c = 0,
                              kernel = 'uniform',
                              cluster = df_terms$cd_municipio_tse)

summary(rdrobust_result)

```


```{r}

dependent <- c("noti_total_mandato",
               "noti_fem_mandato",
               "noti_fisica_mandato",
               "noti_fem_fisica_mandato",
               "noti_psico_mandato",
               "noti_fem_psico_mandato")


results <- map_df(dependent, ~{rdrobust_result <- rdrobust(df_terms[[.]], df_terms$mg_fem,
                              covs = controls,
                              c = 0,
                              kernel = 'uniform',
                              cluster = df_terms$cd_municipio_tse)

  data.frame(
    dependent_var = .,
    band_w = rdrobust_result$bws[1],
    coef = rdrobust_result$coef[3],
    p_value = rdrobust_result$pv[3],
    n_obs_mas = rdrobust_result$N_h[1],
    n_obs_fem = rdrobust_result$N_h[2])})

  

results |> gt()


```

# ROBUSTEZ
## BETA e Pvalor na QUEBRA
```{r}

bandwidths <- seq(.05, 1, by = .1)

results_df <- data.frame(bandwidth = numeric(), p_value = numeric(), beta = numeric())

for (bw in bandwidths) {
  rd_robust_result <- rdrobust(df_terms$tx_noti_total_mandato, df_terms$mg_fem,
                               covs = controls,
                               c = 0,
                               kernel = 'uniform',
                               cluster = df_terms$cd_municipio_tse,
                               h = bw)


  pv <- rd_robust_result$pv[3]
  beta <- rd_robust_result$coef[3]
  
  results_df <- rbind(results_df, data.frame(bandwidth = bw, p_value = pv, beta = beta))
  
}
results_df |> 
  ggplot(aes(y = p_value, x = bandwidth)) + 
  geom_line() + 
  labs(title = "P-valor vs. Largura de Banda",
       x = "Largura de Banda",
       y = "P-valor")

results_df |> 
  ggplot(aes(y = beta, x = bandwidth)) + 
  geom_line() + 
  labs(title = "beta vs. Largura de Banda",
       x = "Largura de Banda",
       y = "beta")


```

```{r}



```


